---
title: "MDA 9169A Project"
author: "Huicong(Ivy) Wu, Moustafa Shaker"
Student ID: "250843449,"
output: 
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read Data
```{r}
#house price data
data = read.csv("https://raw.githubusercontent.com/Waichung1015/Exploratory-Analysis-of-House-Price/master/kc_house_subset.csv")
#predict the house sales in King County, Washington State, USA
#consisted of historic data of houses sold between May 2014 to May 2015. 
#understand which factors are responsible for higher property value - $650K and above
```

1. Data Description
```{r}
library(dplyr)
dim(data)#dimension of dataset
str(data)#data type for each variables
```
2.Data information
```{r}
library(funModeling) 
library(tidyverse) 
library(Hmisc)
#exclude id
df <-data[,-c(1)]
data_stats <- df_status(data)#check data unique, non-zeros, zeros, type and NA values.
#plots the distribution of every numerical variable while automatically excluding the non-numerical ones
# order data by percentages of zeros
arrange(data_stats, -p_zeros) %>% select(variable, q_zeros, p_zeros)


```
Waterfront can be categorical variable.
```{r}
# high carinality variables
#one in which it can take many different values
arrange(data_stats, unique) %>%select(variable, unique)# frequency of lowest rank of variable
freq(df$bedrooms)
freq(df$grade)
freq(df$floors)
freq(df$view)
freq(df$condition)
freq(df$waterfront)
freq(df$bathrooms)
```

3.Data Prepocessing 
3.1 Handle variable"date" format
```{r}
df$date = substr(df$date, 1, 6)
# Converting it to numeric as we can only use numeric values for corrleation
df$date = as.numeric(as.character(df$date))

```

3.2 Remove rows which bathroom ==0 and bedroom ==0
```{r}
df <-df[df$bathrooms > 0,]
df <-df[df$bedrooms > 0,]
```

3.3 Idea 1 for zipcode, add labels
```{r}
len<-nrow(df)

zip_code_trans <-select(df,price,zipcode)
zip_type<-c(1:len)

counter =1
for (i in zip_code_trans$price){
  if (i< 210712 ){zip_type[counter]=0}#<0.05
  if (i>= 210712 && i<249995){zip_type[counter]=1}#0.05
  if (i>= 249995 && i<330000){zip_type[counter]=2}#0.1
  if (i>= 330000 && i<461500){zip_type[counter]=3}#0.25
  if (i>= 461500 && i<660000){zip_type[counter]=4}#0.5
  if (i>= 660000 && i<920150){zip_type[counter]=5}#0.75
  if (i>= 660000 && i<920150){zip_type[counter]=6}#0.90
  if (i>= 920150 && i<1185000){zip_type[counter]=7}#0.95
  if (i>=1185000){zip_type[counter]=8}
  counter=counter +1
}
df["zip_type"]<-zip_type
```
```{r}
boxplot(df$price~df$zip_type)
```
Idea 2 for zipcode, add dummy variables
```{r}
zip_ = factor(df$zipcode)
dummies = model.matrix(~zip_)
df <-cbind(df,dummies[,-c(1)])
```
3.4 Summary of dataset
```{r}
#information about continuous variables, including quantiles, mean, median, count, missing values.
exclude_ <-c("waterfront", "date","zipcode")
exclude_df<-df[ , !(names(df) %in% exclude_)]
describe(exclude_df[-c(18:86)])
```

4. Exploratory Data Analysis
4.1 Pairplots among all variables
```{r}
pairs(price~bedrooms+bathrooms+sqft_living+floors,data = df,col = "blue",pch = 16)
pairs(price~sqft_lot+sqft_above+sqft_basement+sqft_living+sqft_living15+sqft_lot15,data = df,col = "blue",pch = 16)
pairs(price~waterfront+condition+view+grade,data = df,col = "blue",pch = 16)#categorical variables
pairs(price~yr_built+yr_renovated+zipcode,data = df,col = "blue",pch = 16)
pairs(price~lat+long,data = df,col = "blue",pch = 16)
pairs(price~grade+sqft_living+view+bathrooms,data = df,col = "blue",pch = 16)
```
4.2 Analyze response variable - house price distribution
```{r}
#install.packages("ggpubr")
library(ggpubr)
# Basic density plot with mean line and marginal rug
a <- ggplot(df, aes(x = price))
a + geom_histogram(aes(y = ..density..), 
                   colour="black", fill="white",bins=30) +
  geom_density(alpha = 0.2, fill = "#FF6666") 
boxplot(df$price)

```
4.3 Date vs. Price 
```{r}
boxplot(df$price~df$date)

```

4.4 Analyze how age and yr_renew of houses vs. price
```{r}
last_renew<-1:nrow(df)
index =1
df$start = substr(df$date, 1, 4)
df$age<-as.numeric(as.character(df$start))-df$yr_built
for (i in df$yr_renovated){
  #print(i)
  if (i != 0){
     # print(inedx)
      last_renew[index]=as.numeric(df$start)[index]-i
  }
  else{last_renew[index]=0}
  index=index+1
}
print(last_renew)
df$last_renew<-last_renew
plot(df$price~df$age)
plot(df$price~df$last_renew)
```

4.5 Find bathrooms relation with house price
```{r}
price_bath<- select(df,price,bathrooms)
boxplot(price_bath$price~price_bath$bathrooms,xlab="bathrooms",ylab="House Price",main="bathrooms vs. house price",col = "blue")

```

4.6 Find how sqft_living related to house price
```{r}
price_area<- select(df,price,sqft_living)
plot(price_area$price~price_area$sqft_living,xlab="sqft_living",ylab="House Price",main="sqft_living vs. house price",pch =16,col ="blue")

```

4.7 Find how sqft_above related to house price
```{r}
price_above<- select(df,price,sqft_above)
plot(price_above$price~price_above$sqft_above,xlab="sqft_above",ylab="House Price",main="sqft_above vs. house price",pch=16,col ="blue")
```
4.8 Find how sqft_basement related to house price
```{r}
price_above<- select(df,price,sqft_basement)
plot(price_above$price~price_above$sqft_basement,xlab="sqft_basement",ylab="House Price",main="sqft_basement vs. house price",pch=16,col ="blue")
```
4.9 Find how sqft_living15 related to house price
```{r}
price_sqft15<- select(df,price,sqft_living15)
plot(price_sqft15$price~price_sqft15$sqft_living15,xlab="sqft_living",ylab="House Price",main="sqft_living15 vs. house price",pch=16,col ="blue")
```

4.10 lat and long to center
```{r}
new<-paste(df$lat,sep=",",df$long)

distance <-1:nrow(df)
dist_to_center <- function (lat1, lon1, lat2, lon2)
{   
    lon2 <- lon2*pi/180;
    lat2 <- lat2*pi/180;
    lon1 <- lon1*pi/180;
    lat1 <- lat1*pi/180;

    dlon <- lon2 - lon1
    dlat<- lat2 - lat1
    a <- sin(dlat/2)^2 + cos(lat1) * cos(lat2) * sin(dlon/2)^2
    c <- 2 * atan2( sqrt(a), sqrt(1-a) ) 
    d <- 3961 * c;
} 

#print(dist_to_center(47.3487,-122.036,47.6062, 122.3321))      
for (i in 1:nrow(df)){
   distance[i]<-dist_to_center(df$long[i],df$lat[i], -122.200676,47.610378)*1.6
}
df["distance_to_center"]<-distance
plot(df$price~df$distance_to_center)
#min(df$distance_to_center)
#max(df$distance_to_center)

```

4.11 Find how grade affects the house price
```{r}
price_grade<- select(df,price,grade)
boxplot(price_grade$price~price_grade$grade,xlab="grade",ylab="House Price",main="grade vs. house price",col =c("yellow","lightyellow","blue","red","green","orange","purple","pink","grey") ,border = "brown")
abline(h =mean(price_grade$price),col="red",lwd=3, lty=2)
plot(price_grade$price~price_grade$grade,xlab="grade",ylab="House Price",main=" Scatterplot of grade vs. house price")
```

4.12 Find relationship with View and price
```{r}
price_view<- select(df,price,view)
boxplot(price_view$price~price_view$view,xlab="view",ylab="House Price",main="view vs. house price",col = "blue")
plot(price_view$price~price_view$view,xlab="view",ylab="House Price",main="Scatterplot of view vs. house price")

```

4.13 Find relationship with condition and price
```{r}
price_condition<- select(df,price,condition)
boxplot(price_condition$price~price_condition$condition,xlab="condition",ylab="House Price",main="condition vs. house price",col = "blue")
plot(price_condition$price~price_condition$condition,xlab="condition",ylab="House Price",main=" Scatterplot of condition vs. house price")
```
4.14 Find relationship with condition and price
```{r}
price_waterfront<- select(df,price,waterfront)
boxplot(price_waterfront$price~price_waterfront$waterfront,xlab="waterfront",ylab="House Price",main="waterfront vs. house price",col = "blue")

```
4.15 Find relationship with floors and price
```{r}
price_floors<- select(df,price,floors)
boxplot(price_floors$price~price_floors$floors,xlab="floors",ylab="House Price",main="floors vs. house price",pch =16,col ="blue")
```

4.16 Ratio between bathroom and bedrooms
```{r}
bed_to_bath<-df$bedrooms/df$bathrooms
df<-cbind(df,bed_to_bath)
plot(df$price~df$bed_to_bath)
```
4.17 ratio between living_to_lot
```{r}
living_to_lot<-df$sqft_living/df$sqft_lot
df<-cbind(df,living_to_lot)
plot(df$price~df$living_to_lot)
```
4.18 price/sqrt and price/interior
```{r}
price_per_sqrt<-df$price/df$sqft_lot
price_per_interior<-df$price/df$sqft_living
price_per_floor<-df$price/df$floors
df<-cbind(df,price_per_sqrt)
df<-cbind(df,price_per_interior)
df<-cbind(df,price_per_floor)
plot(df$price~df$price_per_sqrt)
plot(df$price~df$price_per_interior)
plot(df$price~df$price_per_floor)
```

4.19 Correlation Matrix and Correlation Heatmap
```{r}
library(ggplot2)
library(reshape2)
#rebuild the dataframe
drop <- c("date","waterfront","zipcode","zip_type","start","yr_renovated","yr_built")
tmp<-df[!(names(df) %in% drop)]
#tmp<-tmp[,(sort(colnames(tmp), decreasing = F))]
grep("zip_98199", colnames(tmp))
df_new<-tmp[-c(16:86)]
corr_matrix_data <- cor(df_new)
#plot correlation matrix
head(corr_matrix_data)#linear dependence between two variables
print("rank of correlation index")
sort(corr_matrix_data[,1])
#default is "pearson"
#plot heatmap
melted_corr_matrix <- melt(corr_matrix_data)
ggplot(data = melted_corr_matrix, aes(x=Var1, y=Var2, fill=value)) + geom_tile() +
theme(text = element_text(size=12),axis.text.x = element_text(angle=90, hjust=1),plot.title = element_text(hjust=0.5))+ggtitle("Correlation Heatmap")+labs( fill = expression("Relation"))+scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1),name="Pearson\nCorrelation") +geom_text(aes(Var2, Var1, label = round(value,2)), color = "black", size = 2)
plot_num(df_new)
```

grade(0.63),sqft-living(0.67),sqft-living15(0.59),sqft_above(0.58),bathrom(0.51)
bathroom,bedroom(0.51); sqft-living,sqft-living15(0.78);sqft_lot,sqft_lot15(0.67);waterfront, view(0.5); grade floor(0.5);
#sqft-living(0.67),sqft_above(0.58),bathrooms(0.51),view(0.5)
#sqft-above should be removed highly correlated to sqft-living,
grade(0.63),sqft-living(0.67),bathrooms(0.51), should be picked.
condition is derived from grade and year-built.

print  correlation values higher than 0.5
```{r}
zdf <- as.data.frame(as.table(cor(df_new)))
subset(zdf, abs(Freq) > 0.5)
```
```{r}
df_new<-df[-c(22:90)]
df_new
```

```{r}
drop <- c("date","zipcode","zip_type","start","sqft_basement","yr_renovated","yr_built","lat","long")
df_new<-df_new[!(names(df_new) %in% drop)]
df_new
write.csv(df_new,'data.csv')
```
```{r}
X =model.matrix(price~.,df_new)[,-1]
y =df_new$price
set.seed (10) 
n = nrow(df_new)
a<-sample(n ,round(0.6*n) , rep=FALSE) 
y_tr <- y[a]
X_tr <- X[a,]

y_ts <- y[-a]
X_ts <- X[-a,]
#train<-df_new[a,] 
#test<-df_new[-a,]
# fit linear model
fit_model = lm(price~., data =df_new[a,] )
summary(fit_model)
pred_model = predict(fit_model, newdata=df_new[-a,]) # prediction for test data
mse_model = mean((pred_model-y_ts)^2) # mse for test data
sqrt(mse_model)
```
```{r}
library("glmnet")
fit_ridge_cv = cv.glmnet(X_tr, y_tr, alpha = 0)
bestlam = fit_ridge_cv$lambda.min
bestlam
plot(fit_ridge_cv)
abline(v=log(bestlam))
fit_ridge_best = glmnet(X_tr, y_tr, alpha = 0, lambda = bestlam)
coef(fit_ridge_best)
round(cbind(coef(fit_model),coef(fit_ridge_best)),2)
pred_ridge = predict(fit_ridge_best, s = bestlam, newx = X_ts)
mse_ridge = mean((pred_ridge-y_ts)^2)
print("MSE for ridge regression:")
print(mse_ridge)
library(caret)
RMSE(pred_ridge, y_ts)
```

```{r}
fit_lasso_cv = cv.glmnet(X_tr, y_tr, alpha = 1) 
bestlam = fit_lasso_cv$lambda.min 
bestlam
plot(fit_lasso_cv)
abline(v=log(bestlam))
fit_lasso_best = glmnet(X_tr, y_tr, alpha = 1, lambda = bestlam)
coef(fit_lasso_best)
round(cbind(coef(fit_model),coef(fit_lasso_best)),2)
pred_lasso = predict(fit_lasso_best,s = bestlam, newx = X_ts) 
mse_lasso = mean((pred_lasso-y_ts)^2)
print("MSE for lasso regression:")
print((mse_lasso))
library(caret)
RMSE(pred_lasso, y_ts)
round(cbind(coef(fit_model),coef(fit_ridge_best),coef(fit_lasso_best)),3)

```

```{r}
c(mse_model,mse_ridge,mse_lasso)
c(RMSE(pred_ridge, y_ts),RMSE(pred_lasso, y_ts))
```





